---
title: "PS2-DanlingMa"
author: "DanlingMa"
date: "2/1/2019"
output: html_document
---
```{r}
rm(list=ls())
library(forecast)
library(ggplot2)
setwd("~/Desktop/Forecasting")
# Passengers (All Carriers - Boston, MA: Logan International (Origin Airport))  
# (www.bts.dot.gov)
logan.data <- read.csv("LoganPassengers.csv")
```
# 1 For this problem load the Logan airport data set from the lecture csv file.
  - Create time series for the domestic and international passengers.
  - Plot both of these lines in the same plot. (Hint, use the plot command, and then the lines command for this. Type help(lines) for help on the lines command.
  - Describe any similarities and differences between the two.
  - Repeat this plot for a short window from 2014 to the end of the data set.
```{r}
# data comes as monthly entries with annual totals (see website)
# need to remove annual values (these are just much bigger than monthly)
# set 
logan.data <- logan.data[logan.data$total<5000000,] 
logan.data.dom = as.data.frame(logan.data[,1])
logan.ts.total <- ts(logan.data$total,start=c(2002,10), freq=12) 
logan.ts.dom <- ts(logan.data$domestic,start=c(2002,10), freq=12) 
logan.ts.inter<- ts(logan.data$international,start=c(2002,10), freq=12)

plot(logan.ts.dom,xlab="Year",ylab="Passengers",bty="l",ylim=range(100000,1500000),col="red")
lines(logan.ts.inter,col="blue")
lines(logan.ts.total,col='green')
legend("topleft",legend=c("Domestic", "International","Total"),col=c("red","blue","green"),lty=1:1, cex=0.8)
grid()
```

